Clustering, profiling and plots

### 2025-01-30 data/02-access-and-prelim-plots.qmd
R1 submitted.

```{r}
load("data-raw/parcels/R_PARCELS_CENTS_2020.rda") #add the census id to the parcels, then the ttm

#first join it to census. Transfer the GeoUID from the DAs that intersect with the residential land use parcel centroids to the R_PARCELS_CENTS_2020 sf object. (we use 'over' here instead of st_intersection because it is way faster)
R_PARCELS_CENTS_GeoUIDs <- sp::over(as_Spatial(R_PARCELS_CENTS_2020),as_Spatial(HAM_census_21 |> dplyr::select(GeoUID)))

@@

```{r, eval=FALSE}
care_categories_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.col = "black", lwd=1)+
  tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'), size=0.025, title = "")+
  tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("left", "top"), size=0.9) +
  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 1, legend.bg.color = "white", legend.bg.alpha = 0.4,
            panel.labels = c("1,265 Dependent-centric destinations", 
                             "158 Errand-centric destinations", 
                             "381 Grocery-centric destinations", 
                             "421 Health-centric destinations"),
            panel.label.size = 1.1)

tmap_save(care_categories_plot, 
          file= file.path(here("figures/care_categories_plot.png")),
          dpi=300)
```

>>>>

```{r}
load("data-raw/parcels/R_PARCELS_2020.rda") #add the census id to the parcels, then the ttm
R_PARCELS_CENTS_2020 <- st_centroid(R_PARCELS_2020)
#first join it to census. Transfer the GeoUID from the DAs that intersect with the residential land use parcel centroids to the R_PARCELS_CENTS_2020 sf object. (we use 'over' here instead of st_intersection because it is way faster)
R_PARCELS_CENTS_GeoUIDs <- sp::over(as_Spatial(R_PARCELS_CENTS_2020),as_Spatial(HAM_census_21 |> dplyr::select(GeoUID)))

@@

```{r, eval=FALSE}
care_categories_plot <- 
  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.alpha=0)+
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+ 
    tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = NA, alpha=0, border.col = "black", lwd=1)+
  tm_shape(care_dest) + tm_symbols(shape="Care_Category_Specific", size=0.03, border.lwd = 0, shapes = c(21,22,23,24,25,15), col = "#E6AB02", title.shape = "")+
  # tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'),
  #                              size=0.025, title = "")+
  tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+ 
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("left", "top"), size=0.9) +
  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), scale = 1.1, legend.bg.color = "white", legend.bg.alpha = 0.65,
            panel.labels = c("1,265 Dependent-centric destinations", 
                             "158 Errand-centric destinations", 
                             "381 Grocery-centric destinations", 
                             "421 Health-centric destinations"),
            panel.label.size = 1.1) 
#care_categories_plot
tmap_save(care_categories_plot,
          file= file.path(here("figures/care_categories_plot.png")),
          dpi=600)
```

### 2025-01-30 data/03-clustering-and-profiling.qmd
R1 submitted.

```{r, eval=FALSE}
INCOME_B_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("Income_bottom",
              border.alpha = 0.2,
              palette = "Greens", style = "quantile", n=4,
              title = "% hld below AT median income",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

AGE014_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("Age_0to14",
              border.alpha = 0.2,
              palette = "Oranges", style = "quantile", n=4,
              title = "% population aged 0 to 14",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

ONE_PARENT_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("One_parent_families",
              border.alpha = 0.2,
              palette = "Reds", style = "quantile", n=4,
              title = "% one parent hlds",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

LICOAT_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("LICO",
              border.alpha = 0.2,
              palette = "Purples", style = "quantile", n=4,
              title = "Prevalance of household LICO-AT",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

@@

```{r, eval=FALSE}
dep_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("med_dep_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Oranges",
              title = "Dependent-centric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

health_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("med_health_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Purples",
              title = "Health-centric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

groc_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("med_groc_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Greens",
              title = "Grocery-centric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

err_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("med_err_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Reds",
              title = "Errand-centric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

@@

```{r, eval=FALSE}
all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("med_div",
              border.alpha = 0.2,
              style = "quantile", n=3, #are the quantiles, but are also similar to the clustered profiles
              label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
              palette = "Greys",
              title = "Diversity of care destinations within a 15-minute walk for a median parcel",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

>>>>

```{r, eval=FALSE}
INCOME_B_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("Income_bottom",
              border.alpha = 0.2,
              palette = "Greens", style = "quantile", n=4,
              title = "% hld below AT median income",
              legend.is.portrait = TRUE, 
              colorNA = "white",
              textNA= "Unavailable") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

AGE014_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("Age_0to14",
              border.alpha = 0.2,
              palette = "Oranges", style = "quantile", n=4,
              title = "% population aged 0 to 14",
              legend.is.portrait = TRUE, 
              colorNA = "white",
              textNA= "Unavailable") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

ONE_PARENT_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("One_parent_families",
              border.alpha = 0.2,
              palette = "Reds", style = "quantile", n=4,
              title = "% one parent hlds",
              legend.is.portrait = TRUE, 
              colorNA = "white",
              textNA= "Unavailable") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

LICOAT_DA_plot <-
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("LICO",
              border.alpha = 0.2,
              palette = "Purples", style = "quantile", n=4,
              title = "Prevalance of household LICO-AT",
              legend.is.portrait = TRUE, 
              colorNA = "white",
              textNA= "Unavailable") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

@@

```{r, eval=FALSE}
dep_copp_plot <- 
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(med_dep_copp = ifelse(med_dep_copp == 0, NA, med_dep_copp))) +
  tm_polygons("med_dep_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Oranges",
              title = "Dependent-centric",
              colorNA = "white",
              textNA = 0,
              border.col = NULL) +
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, scale=0.8,
            bg.color = "grey",
            legend.position = c("right","top"))

health_copp_plot <- 
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(med_health_copp = ifelse(med_health_copp == 0, NA, med_health_copp))) +
  tm_polygons("med_health_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Purples",
              title = "Health-centric",
              colorNA = "white",
              textNA = 0,
              border.col = NULL) +
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, scale=0.8,
            bg.color = "grey",
            legend.position = c("right","top"))

groc_copp_plot <-
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(med_groc_copp = ifelse(med_groc_copp == 0, NA, med_groc_copp))) +
  tm_polygons("med_groc_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Greens",
              title = "Grocery-centric",
              colorNA = "white",
              textNA = 0,
              border.col = NULL) +
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, scale=0.8,
            bg.color = "grey",
            legend.position = c("right","top"))

err_copp_plot <- 
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(med_err_copp = ifelse(med_err_copp == 0, NA, med_err_copp))) +
  tm_polygons("med_err_copp",
              border.alpha = 0.2,
              style = "quantile", n=4,
              palette = "Reds",
              title = "Errand-centric",
              colorNA = "white",
              textNA = 0,
              border.col = NULL) +
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, scale=0.8,
            bg.color = "grey",
            legend.position = c("right","top"))

@@

```{r, eval=FALSE}
all_cats_entropy_copp_plot <- 
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(med_div = ifelse(med_div == 0, NA, med_div))) +
  tm_polygons("med_div",
              border.alpha = 0.2,
              style = "quantile", n=4, #are the quantiles, but are also similar to the clustered profiles
              #label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
              palette = "Greys",
              title = "Diversity of care destinations within \na 15-minute walk for a median parcel",
              legend.is.portrait = TRUE,
              colorNA = "white",
              textNA = 0,
              border.col = NULL) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            #legend.outside = TRUE,
            legend.position = c("right", "top"))
            #legend.outside.position = "bottom",
            #legend.outside.size = 0.2,
            # legend.width = 12)
